[dcpo_demsupport]

### 2020-06-12 paper/dcpo_demsupport.Rmd
add note and fix per @Tyhcass re claassen_input_raw1; paste in AJPS (not PA) replication script (still messy at the moment)

```{r claassen_input}
claassen_input_raw1 <- read_csv("data/claassen_input_raw.csv") %>% 
  filter(!(str_detect(item, "army") & # WVS obs identified as problematic by Claassen 
    ((country=="Albania" & year==1998) |
         (country=="Indonesia" & (year==2001 | year==2006)) |
         (country=="Iran" & year==2000) |
         (country=="Pakistan" & (year==1996 | year==2001)) |
         (country=="Vietnam" & year==2001)) |
    (str_detect(item, "strong") &
         ((country=="Egypt" & year==2012) |
              (country=="Iran" & (year==2000 | year==2005)) |
              (country=="India") |
              (country=="Pakistan" & (year==1996 | year==2001)) |
              (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
              (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
              (country=="Vietnam" & year==2001))))) %>% 
  filter(!(survey=="evs2017" | survey == "arabb5" | survey == "eb47")) %>%  # surveys unused by Claassen
  mutate(item = if_else(item == "strong_amb_1" & year == 2004, "strong_amb_2", item)) # items conflated in amb_combo file

@@

supdem <- get_file("supdem raw survey marginals.tab", "doi:10.7910/DVN/HWLW0J") %>% 
  read_csv(col_types = cols(.default = col_double(), 
                            Country = col_character(),
                            CAbb = col_character(),
                            Project = col_character()))

@@

sddem1 = gather(data=supdem, key="Item", value="Response", 6:33, na.rm=TRUE) 
sddem1$Sample = 0
sddem1[sddem1$Project=="AfB", "Sample"] = sddem1[sddem1$Project=="AfB", "AfB_sample"]
sddem1[sddem1$Project=="ArB", "Sample"] = sddem1[sddem1$Project=="ArB", "ArB_sample"]
sddem1[sddem1$Project=="AsB", "Sample"] = sddem1[sddem1$Project=="AsB", "AsB_sample"]
sddem1[sddem1$Project=="AsnB", "Sample"] = sddem1[sddem1$Project=="AsnB", "AsnB_sample"]
sddem1[sddem1$Project=="ESS", "Sample"] = sddem1[sddem1$Project=="ESS", "ESS_sample"]
sddem1[sddem1$Project=="LAPOP", "Sample"] = sddem1[sddem1$Project=="LAPOP", "LAPOP_sample"]
sddem1[sddem1$Project=="LB", "Sample"] = sddem1[sddem1$Project=="LB", "LB_sample"]
sddem1[sddem1$Project=="NDB", "Sample"] = sddem1[sddem1$Project=="NDB", "NDB_sample"]
sddem1[sddem1$Project=="Pew", "Sample"] = sddem1[sddem1$Project=="Pew", "Pew_sample"]
sddem1[sddem1$Project=="WVS", "Sample"] = sddem1[sddem1$Project=="WVS", "WVS_sample"]
sddem1[sddem1$Project=="CSES", "Sample"] = sddem1[sddem1$Project=="CSES", "CSES_sample"]

# trim stacked dataset
sddem1 = sddem1[, -c(6:16)]
sddem1$Response = ifelse(sddem1$Response==0, NA, sddem1$Response)
sddem1 = sddem1[!is.na(sddem1$Response), ]

# create item by project indicators
sddem1 = unite(sddem1, ItemProj, c(Project, Item), sep = "_", remove = FALSE)
sddem1$ItemProj = as.factor(sddem1$ItemProj)

# create year by country indicators
year0 = 1990
sddem1 = sddem1[sddem1$Year > year0,]
sddem1 = unite(sddem1, YearCountry, c(Year, Country), sep = "_", remove = FALSE)
sddem1$YearCountry = as.factor(sddem1$YearCountry)

# create item by project by country indicators
sddem1 = unite(sddem1, ItemProjCnt, c(ItemProj, Country), sep = "_", remove = FALSE)
sddem1$ItemProjCnt = as.factor(sddem1$ItemProjCnt)

# create item by country indicators
sddem1 = unite(sddem1, ItemCnt, c(Item, Country), sep = "_", remove = FALSE)
sddem1$ItemCnt = as.factor(sddem1$ItemCnt)

# create year by project indicators
sddem1 = unite(sddem1, YrProj, c(Year, Project), sep = "_", remove = FALSE)
sddem1$YrProj = as.factor(sddem1$YrProj)

# create year by project by country indicators
sddem1 = unite(sddem1, YrProjCnt, c(YrProj, Country), sep = "_", remove = FALSE)
sddem1$YrProjCnt = as.factor(sddem1$YrProjCnt)

# factorise
sddem1$Country = as.factor(as.character(sddem1$Country))
sddem1$Item = as.factor(as.character(sddem1$Item))
sddem1$ItemProj = as.factor(as.character(sddem1$ItemProj))
sddem1$ItemCnt = as.factor(as.character(sddem1$ItemCnt))
sddem1$ItemProjCnt = as.factor(as.character(sddem1$ItemProjCnt))
sddem1$Project = as.factor(as.character(sddem1$Project))
sddem1$YrProj = as.factor(as.character(sddem1$YrProj))
sddem1$YrProjCnt = as.factor(as.character(sddem1$YrProjCnt))
sddem1$Year = as.factor(sddem1$Year-year0)

# drop countries with less than 2 years of data
cnt.obs.years = rowSums(table(sddem1$Country, sddem1$Year) > 0)
sddem2 = sddem1[sddem1$Country %in% levels(sddem1$Country)[cnt.obs.years > 1], ]

## Stan estimation

# prepare data for stan

n.items = length(unique(sddem2$Item))
n.cntrys = length(unique(sddem2$Country))
n.yrs = 2015-year0
n.proj = length(unique(sddem2$Project))
n.resp = dim(sddem2)[1]
n.itm.prj = length(unique(sddem2$ItemProj))
n.itm.cnt = length(unique(sddem2$ItemCnt))
n.cntry.yrs = n.cntrys * n.yrs
n.itm.prj.cnt = length(unique(sddem2$ItemProjCnt))
n.yr.proj.cnt = length(unique(sddem2$YrProjCnt))
cntrys = as.numeric(factor(sddem2$Country))
cnt.names = sort(unique(sddem2$Country))
cnt.ccode = sddem2[match(cnt.names, sddem2$Country), "COWCode"]
items = as.numeric(factor(sddem2$Item))
yrs = as.numeric(sddem2$Year)
projs = as.numeric(factor(sddem2$Project))
itm.prjs = as.numeric(factor(sddem2$ItemProj))
itm.prj.cnts = as.numeric(factor(sddem2$ItemProjCnt))
itm.cnts = as.numeric(factor(sddem2$ItemCnt))
cntry.yrs = as.numeric(sddem2$YearCountry)

# specify data for stan
dat.2 = list(N=n.resp, K=n.itm.prj, T=n.yrs, J=n.cntrys, P=n.itm.prj.cnt, jj=cntrys, tt=yrs, 
             pp=itm.prj.cnts, kk=itm.prjs, x=sddem2$Response, samp=sddem2$Sample)

# pars
pars.5 = c("mu_lambda","sigma_lambda","sigma_delta","sigma_theta","phi","lambda","delta","theta","x_pred","log_lik")

# set pars for final run (time-consuming; reduce iterations for trial run)
n.iter = 1000
n.warm = 500
n.chn = 4
n.thin = 2

## supdem.stan.mod5.stan obtained from dataverse replication files
stan.mod.5 <- stan(file = "supdem.stan.mod5.stan", data=dat.2, pars=pars.5, 
                   iter=n.iter, warmup=n.warm, chains=n.chn, thin=n.thin, 
                   control=list(adapt_delta=0.99, stepsize=0.02, max_treedepth=11))
```

>>>>

```{r claassen_input}
claassen_input_raw1 <- read_csv("data/claassen_input_raw.csv") %>% 
  filter(!(str_detect(item, "army") & # WVS obs identified as problematic by Claassen (note: these are his years right now)
    ((country=="Albania" & year==1998) |
         (country=="Indonesia" & (year==2001 | year==2006)) |
         (country=="Iran" & year==2000) |
         (country=="Pakistan" & (year==1996 | year==2001)) |
         (country=="Vietnam" & year==2001)) |
    (str_detect(item, "strong") &
         ((country=="Egypt" & year==2012) |
              (country=="Iran" & (year==2000 | year==2005)) |
              (country=="India") |
              (country=="Pakistan" & (year==1996 | year==2001)) |
              (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
              (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
              (country=="Vietnam" & year==2001))))) %>% 
  filter(!(survey=="evs2017" | survey == "arabb5" | survey == "eb471")) %>%  # surveys unused by Claassen
  mutate(item = if_else(item == "strong_amb_1" & year == 2004, "strong_amb_2", item)) # items conflated in amb_combo file

@@

supdem <- get_file("supdem raw survey marginals.tab", "doi:10.7910/DVN/HWLW0J") %>% 
  read_csv(col_types = "cdcddcdc")

@@

library(MASS)
library(arm)
library(dplyr)
library(tidyr)
library(rstan)
library(loo)
library(ggplot2)
library(RColorBrewer)

# options
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# working directory
WD = getwd()
setwd(WD)

# read support for dem data
sddem1 = read.csv("supdem raw survey marginals.csv")

# remove NAs
sddem1 = sddem1[!sddem1$Response==0, ]

# create year by country indicators
year0 = 1987
sddem1 = sddem1[sddem1$Year > year0,]
sddem1 = unite(sddem1, YearCountry, c(Year, Country), sep = "_", remove = FALSE)

# create item by country indicators
sddem1 = unite(sddem1, ItemCnt, c(Item, Country), sep = "_", remove = FALSE)

# create year by project indicators
sddem1 = unite(sddem1, YrProj, c(Year, Project), sep = "_", remove = FALSE)

# create year by project by country indicators
sddem1 = unite(sddem1, YrProjCnt, c(YrProj, Country), sep = "_", remove = FALSE)

# factorise
sddem1$Country = as.factor(as.character(sddem1$Country))
sddem1$Item = as.factor(as.character(sddem1$Item))
sddem1$ItemCnt = as.factor(as.character(sddem1$ItemCnt))
sddem1$Project = as.factor(as.character(sddem1$Project))
sddem1$YrProj = as.factor(as.character(sddem1$YrProj))
sddem1$YrProjCnt = as.factor(as.character(sddem1$YrProjCnt))
sddem1$Year = sddem1$Year-year0

# count data
dim(sddem1)[1]                   # 3765 national responses
length(unique(sddem1$Country))   # 150 countries
length(unique(sddem1$Project))   # 14 projects
length(unique(sddem1$Item))      # 52 items
length(unique(sddem1$ItemCnt))   # 1453 item-country
length(unique(sddem1$YrProjCnt)) # 1390 national surveys
length(unique(sddem1$Year))      # 27 unique years (out of 30)
sum(sddem1$Sample) / dim(sddem1)[1] * length(unique(sddem1$YrProjCnt))   # 1,804,450 respondents

# drop countries with less than 2 years of data
cnt.obs.years = rowSums(table(sddem1$Country, sddem1$Year) > 0)
sort(cnt.obs.years)
sddem2 = sddem1[sddem1$Country %in% levels(sddem1$Country)[cnt.obs.years > 1], ]
length(unique(sddem2$Country))   # 137 countries with 2+ years of data

#### Stan estimation

# prepare data for stan
n.items = length(unique(sddem2$Item))
n.cntrys = length(unique(sddem2$Country))
n.yrs = 2017-year0
n.proj = length(unique(sddem2$Project))
n.resp = dim(sddem2)[1]
n.itm.cnt = length(unique(sddem2$ItemCnt))
n.cntry.yrs = n.cntrys * n.yrs
n.yr.proj.cnt = length(unique(sddem2$YrProjCnt))
cntrys = as.numeric(factor(sddem2$Country))
cnt.names = as.character(sort(unique(sddem2$Country)))
cnt.code = as.character(sddem2[match(cnt.names, sddem2$Country), "CAbb"])
cnt.code[83] = "MNE"
cnt.code[106] = "SRB"
items = as.numeric(factor(sddem2$Item))
yrs = sddem2$Year
projs = as.numeric(factor(sddem2$Project))
itm.cnts = as.numeric(factor(sddem2$ItemCnt))

# specify data for stan
dat.1 = list(N=n.resp, K=n.items, T=n.yrs, J=n.cntrys, jj=cntrys, tt=yrs, kk=items, 
             x=sddem2$Response, samp=sddem2$Sample)
dat.2 = list(N=n.resp, K=n.items, T=n.yrs, J=n.cntrys, P=n.itm.cnt, jj=cntrys, tt=yrs, 
             pp=itm.cnts, kk=items, x=sddem2$Response, samp=sddem2$Sample)
sapply(dat.2, summary)

# pars
pars.4 = c("mu_lambda","sigma_lambda","sigma_theta","phi","lambda","theta","x_pred","log_lik")
pars.5 = c("mu_lambda","sigma_lambda","sigma_delta","sigma_theta","phi","lambda","delta","theta",
            "x_pred","log_lik")
pars.6 = c("Sigma","Omega","sigma_delta","sigma_theta","phi","lambda","gamm","delta","theta",
           "x_pred","log_lik")

# iterations
n.iter = 500
n.warm = 250
n.chn = 4
n.thin = 1

# best model from PA article (model 5)
stan.mod.5 = stan(file='supdem.stan.mod5.stan', data=dat.2, pars=pars.5, 
                   iter=n.iter, warmup=n.warm, chains=n.chn, thin=n.thin, 
                   control=list(adapt_delta=0.99, stepsize=0.02, max_treedepth=11))

# model 5 with different priors
stan.mod.5.2 = stan(file='supdem.stan.mod5.2.stan', data=dat.2, pars=pars.5, 
                  iter=n.iter, warmup=n.warm, chains=n.chn, thin=n.thin, 
                  control=list(adapt_delta=0.99, stepsize=0.02, max_treedepth=11))

# model 5 with item slopes
stan.mod.6 = stan(file='supdem.stan.mod6.2.stan', data=dat.2, pars=pars.6, 
                   iter=n.iter, warmup=n.warm, chains=n.chn, thin=n.thin, 
                   control=list(adapt_delta=0.99, stepsize=0.01, max_treedepth=13))
```
